---
title: "Hydra OM diagnostics"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      knitr.duplicate.label = "allow")
suppressPackageStartupMessages(library(tidyverse))
source("R/read.report.R")
source("R/gettables.R")
library(tidyverse)
library(hydradata)
library(compResidual)
```


```{r}

#READ SARAHS DATA SETS
hydraDataList <- readRDS("inputs/hydra_sim_GBself_5bin.rds")

#READ FITS FROM SARAHS DATA SETS
repfile <- "inputs/OM/hydra_sim.rep"
#repfile <- "inputs/initial_run/hydra_sim.rep"

output<-reptoRlist(repfile)

#### READ CATCH AND SURVEY OBSERVED BIOMASS ####
obs_surveyB <- hydraDataList$observedBiomass
obs_catchB <- hydraDataList$observedCatch

biorows <- dim(obs_surveyB)[1]
catchrows <- dim(obs_catchB)[1]

#create a table with estimated and observed values
indexfits <- gettables(repfile, biorows, catchrows)
```

#### abundance indices

```{r}
survey_obspred <- indexfits[1][[1]] %>%
  mutate(obs = biomass + 1e-07,
         pred = pred_bio + 1e-07,
         log_obs = log(obs),
         log_pred = log(pred),
         log_lo = log_obs - 1.96*cv,
         log_hi = log_obs + 1.96*cv,
         obs_lo = exp(log_lo),
         obs_hi = exp(log_hi),
         species = hydraDataList$speciesList[species])
```

#### fits to survey indices of abundance {.tabset}

```{r, results = 'asis'}
for (surv in unique(survey_obspred$survey)) {
  cat("##### Survey #", surv, " \n")
  p1 <- survey_obspred %>% 
    filter(survey == surv) %>% 
    ggplot() +
    aes(x= year, y = log_obs, group = species, col=factor(survey)) +
    geom_errorbar(aes(ymin = log_lo, ymax = log_hi)) +
    geom_point() +
    geom_line(aes(x=year, y=log_pred), col = "blue") +
    facet_wrap(~species, scales = "free_y") +
    theme_bw() +
    guides(species = "None")
  print(p1)
  cat(" \n\n")
}

```

#### catch

```{r}
catch_obspred <- indexfits[2][[1]] %>% 
  mutate(obs = catch + 1e-07,
         pred = pred_catch + 1e-07,
         log_obs = log(obs),
         log_pred = log(pred),
         log_lo = log_obs - 1.96*cv,
         log_hi = log_obs + 1.96*cv,
         obs_lo = exp(log_lo),
         obs_hi = exp(log_hi),
         species = hydraDataList$speciesList[species])
```

#### fits to Catch time series {.tabset} 

```{r, results = 'asis'}
for (fleet in unique(catch_obspred$fishery)) {
  cat("##### Fleet #", fleet, " \n")
  p3 <- catch_obspred %>% 
    filter(fishery == fleet,
           catch > 0) %>% 
    ggplot() +
    aes(x= year, y = log_obs, group = species, col= factor(fishery)) +
    geom_errorbar(aes(ymin = log_lo, ymax = log_hi)) +
    geom_point() +
    geom_line(aes(x=year, y=log_pred), col = "blue") +
    facet_wrap(~species, scales = "free_y") +
    theme_bw() +
    guides(species = "None")
  print(p3)
  cat(" \n\n")
}
```

#### {-}

#### Estimated Recruitment time series

```{r}
est_recruits <- output$EstRec %>%
  as.data.frame() %>% 
  mutate(species = hydraDataList$speciesList) %>% 
  select(species, everything()) %>% 
  pivot_longer(cols = -species, names_to = "year") %>% 
  mutate(year = as.integer(str_remove(year, "V")),
         log_rec = ifelse(value > 0,log(value),NA))
```

```{r}
est_recruits %>% 
  ggplot() +
  aes(x = year, y = value) +
  geom_point() +
  geom_line() +
  facet_wrap(~species, scales = "free") +
  theme_minimal() +
  labs(x = "Year",
       y = "Recruitment (thousands)",
       title = "Time series of estimated recruitment")
```
time series of log-recruitment

```{r}
est_recruits %>% 
  ggplot() +
  aes(x = year, y = log_rec) +
  geom_point() +
  geom_line() +
  facet_wrap(~species, scales = "free") +
  theme_minimal() +
  labs(x = "Year",
       y = "ln(Recruitment)",
       title = "Time series of LN(Recruitment)")
```

#### Biomass time series

```{r}

# time steps per year varies by model, need to save it
# updated rep file
stepperyr <- output$Nstepsyr
if (length(stepperyr)==0) stepperyr <- nrow(output$EstBsize)/hydraDataList$Nyrs/length(hydraDataList$speciesList)

est_bio <- output$EstBsize %>%
  rowSums() %>% 
  as.data.frame() %>% 
  rename(bio = ".") %>% 
  mutate(species = rep(hydraDataList$speciesList, each = hydraDataList$Nyrs*stepperyr),
         year  = rep(1:(hydraDataList$Nyrs*stepperyr),hydraDataList$Nspecies),
         year = (1-1/stepperyr) + year / stepperyr,  #5 time steps per year
         log_bio = ifelse(bio>0,log(bio),NA))
```


```{r}
est_bio %>% 
  ggplot() +
  aes(x = year, y = bio) +
  #geom_point() +
  geom_line() +
  facet_wrap(~species, scales = "free") +
  theme_minimal() +
  labs(x = "Year",
       y = "Biomass (t)",
       title = "Time series of estimated biomass")
```


```{r, eval=FALSE}
est_bio %>% 
  ggplot() +
  aes(x = year, y = log_bio) +
  #geom_point() +
  geom_line() +
  facet_wrap(~species, scales = "free") +
  theme_minimal() +
  labs(x = "Year",
       y = "Biomass (t)",
       title = "Time series of estimated LN(biomass)")
```

#### Fishing mortality time series

```{r}
est_F <- output$Fyr %>% 
  as.data.frame() %>% 
  pivot_longer(cols=3:ncol(.), names_to = "year", names_prefix = "V") %>% 
  rename(species = "V1",
         fleet = "V2") %>% 
  mutate(year = as.numeric(year)-2,
         species = hydraDataList$speciesList[species])
```

```{r}
est_F %>% 
  ggplot() +
  aes(x = year, y = value, col = factor(fleet)) +
  #geom_point() +
  geom_line() +
  facet_wrap(~species, scales = "free") +
  theme_minimal() +
  labs(x = "Year",
       y = "F",
       col = "fleet",
       title = "Time series of estimated fishing mortality")
```
```{r}
nlen <- ncol(output$EstM2size)
est_M2 <- output$EstM2size %>% 
  as.data.frame() %>% 
  pivot_longer(cols=1:ncol(.), names_to = "ilen", names_prefix = "V") %>% 
  mutate(species = rep(hydraDataList$speciesList, each = hydraDataList$Nyrs*nlen*stepperyr),
         year  = rep(rep(1:(hydraDataList$Nyrs*stepperyr),each=nlen), length(hydraDataList$speciesList)),
         year = (1-1/stepperyr) + year / stepperyr) %>%
  I()
```

```{r}
est_M2 %>% 
  ggplot() +
  aes(x = year, y = value, col = factor(ilen)) +
  #geom_point() +
  geom_line() +
  facet_wrap(~species, scales = "free") +
  theme_minimal() +
  labs(x = "Year",
       y = "M2",
       col = "length bin",
       title = "Time series of estimated predation mortality") +
  #theme(legend.position = "bottom") +
  #    labs(col="") +
  #    guides(col = guide_legend(nrow = 1))
  NULL
```
#### Survey selectivity

```{r}
survey_sel <- output$survey_sel %>% 
  as.data.frame() %>% 
  pivot_longer(cols = 3:ncol(.), names_prefix = "V", names_to = "ilen") %>% 
  rename(species = "V1",
         survey = "V2") %>% 
  mutate(ilen = as.numeric(ilen) - 2,
         species = hydraDataList$speciesList[species]) %>% 
  I()
```
```{r}
survey_sel %>% 
  ggplot() +
  aes(x=ilen, y = value, col = factor(survey)) +
  geom_line() +
  facet_wrap(~species) +
  labs(x = "length bin",
       y = "selectivity",
       col = "survey") +
  NULL
```

#### Fishery selectivity

```{r}
survey_sel <- output$fishsel %>% 
  as.data.frame() %>% 
  pivot_longer(cols = 3:ncol(.), names_prefix = "V", names_to = "ilen") %>% 
  rename(species = "V1",
         fleet = "V2") %>% 
  mutate(ilen = as.numeric(ilen) - 2,
         species = hydraDataList$speciesList[species]) %>% 
  I()
```
```{r}
survey_sel %>% 
  ggplot() +
  aes(x=ilen, y = value, col = factor(fleet)) +
  geom_line() +
  facet_wrap(~species) +
  labs(x = "length bin",
       y = "selectivity",
       col = "fleet") +
  NULL
```
#### size comps

```{r, echo = FALSE}
obs_survey <- hydraDataList$observedSurvSize %>% tibble()
obs_survey <- obs_survey %>% pivot_longer(cols=6:ncol(.), names_to = "lenbin") %>% #filter(value != -999)%>%
  
  mutate(lenbin = as.integer(str_remove(lenbin, "sizebin")),
         label = rep("survey",nrow(.)),
         species = hydraDataList$speciesList[species])
obs_survey$value[which(obs_survey$value == -999)] = 0.00001

pred_surv<-output$pred_survey_size
obs_survey$pred_surv<-pred_surv
nll_survey<-output$nll_survey_size
obs_survey$nll_surv<-nll_survey
obs_survey$pearson<-((obs_survey$value-obs_survey$pred_surv)/sqrt(obs_survey$pred_surv))

colnames(obs_survey) <- c('number','year','species','type','InpN','lenbin','obs_value','label',
                          'pred_value','nll','pearson')
```


```{r, echo = FALSE}
obs_catch <- hydraDataList$observedCatchSize %>% tibble()
obs_catch<-obs_catch %>% pivot_longer(cols=7:ncol(.), names_to = "lenbin") %>%
  mutate(lenbin = as.integer(str_remove(lenbin, "sizebin")),
         label = rep("catch",nrow(.)),
         species = hydraDataList$speciesList[species])# %>% filter(value != -999)
obs_catch$value[which(obs_catch$value == -999)] = 0.00001

pred_catch<-output$pred_catch_size
obs_catch$pred_catch<-pred_catch
nll_catch<-output$nll_catch_size
obs_catch$nll_catch<-nll_catch
obs_catch$pearson<-((obs_catch$value-obs_catch$pred_catch)/sqrt(obs_catch$pred_catch))

colnames(obs_catch) <- c('number','area','year','species','type','InpN','lenbin','obs_value','label',
                         'pred_value','nll','pearson')

diet_catch <- bind_rows(obs_catch, obs_survey)
```


```{r, echo = FALSE}
obs_diet <- hydraDataList$observedSurvDiet %>% tibble()
obs_diet<-obs_diet %>% pivot_longer(cols=6:ncol(.), names_to = "prey") %>%
  mutate(#lenbin = as.integer(str_remove(lenbin, "V")),
    species = hydraDataList$speciesList[species],
    label = rep("diet",nrow(.))) %>%
  I()


pred_diet<-output$pred_dietprop

if (length(pred_diet)!=nrow(obs_diet)) obs_diet <- obs_diet %>% filter(value != 0)

obs_diet$pred_diet<-pred_diet
nll_diet<-output$nll_dietprop
obs_diet$nll_diet<-nll_diet
obs_diet$pearson<-((obs_diet$value-obs_diet$pred_diet)/sqrt(obs_diet$pred_diet))
colnames(obs_diet) <- c('number','year','species','lenbin','InpN','prey','obs_value','label', 'pred_value','nll','pearson')

mydata <- diet_catch %>%
  mutate(#prey = replace_na(prey, -99),
    area = replace_na(area, 1),
    type = replace_na(type, 0))
#data <- mydata
mydata$residual<-ifelse(mydata$pearson<0,"negative","positive")
mydata$res_abs<-abs(mydata$pearson)
mydata<-as.data.frame(mydata)

temp.catch = mydata[which(mydata$label == "catch"),]
temp.surv = mydata[which(mydata$label == "survey"),]
#temp.diet = data[which(data$label == "diet"),]

mydata <- obs_diet #data <- read.csv("outputs/survdiet.csv", header = T)
mydata$residual<-ifelse(mydata$pearson<0,"negative","positive")
mydata$res_abs<-abs(mydata$pearson)
temp.diet = mydata[which(mydata$label == "diet"),]
```

##### surveys

```{r}
long_surv <- temp.surv %>%
  pivot_longer(cols = c("pred_value","obs_value"),
               names_to = c("kind","junk"),names_sep = "_") %>%
  select(-junk)

sp<-1
plot_surv <- list()
especies<-unique(long_surv$species)
for (sp in especies) {
  
  temp_size<-long_surv %>% filter(species == sp & number==1) %>%
    group_by(year) %>%
    summarize(mu_ss=mean(InpN))
  
  plot_surv[[sp]] <- long_surv %>% filter (species==sp& number==1) %>%
    ggplot() +
    aes(x=lenbin, y = value) +
    geom_line(aes(col = kind)) +
    geom_point(data = long_surv %>% filter (species==sp, number==1, kind == "obs"),
               aes(x=lenbin, y=value)) +
    facet_wrap(~year, dir="v") +
    geom_text(data=temp_size, aes(x = 4.5, y = 0.5, label = mu_ss), size=3) +
    theme(legend.position = "bottom") +
    labs(col="") +
    guides(col = guide_legend(nrow = 1))
  
  #ggsave(paste0(plotdir,"complot_surv_",sp,".jpeg"), plot_surv, width = 10, height = 7, dpi = 300)#, width=3000, height=2500, res=250)
  
}

```

##### Size composition of survey 1 by species {.tabset}
```{r, results = 'asis', fig.height=10, fig.width=9}
for (sp in especies){
  tmp <- plot_surv[[sp]]
  cat("######", sp, " \n")
  print(tmp)
  cat(" \n\n")
}

```
##### {-}

```{r}
sp<-1
plot_surv <- list()
especies<-unique(long_surv$species)
for (sp in especies) {
  
  temp_size<-long_surv %>% filter(species == sp & number==2) %>%
    group_by(year) %>%
    summarize(mu_ss=mean(InpN))
  
  if(dim(temp_size)[1] > 0) {
    
    plot_surv[[sp]] <- long_surv %>% filter (species==sp & number==2) %>%
      ggplot() +
      aes(x=lenbin, y = value) +
      geom_line(aes(col = kind)) +
      geom_point(data = long_surv %>% filter (species==sp, number==2, kind == "obs"),
                 aes(x=lenbin, y=value)) +
      facet_wrap(~year, dir="v", drop = FALSE) +
      geom_text(data=temp_size, aes(x = 4.5, y = 0.5, label = mu_ss), size=3) +
      theme(legend.position = "bottom") +
      labs(col="") +
      guides(col = guide_legend(nrow = 1))
    
  }
  
  if(dim(temp_size)[1] == 0) plot_surv[[sp]] <- NULL
  
  #ggsave(paste0(plotdir,"complot_surv_",sp,".jpeg"), plot_surv, width = 10, height = 7, dpi = 300)#, width=3000, height=2500, res=250)
  
}

```

##### Size composition of survey 2 by species {.tabset}

```{r, results = 'asis', fig.height=10, fig.width=9}
for (sp in especies){
  tmp <- plot_surv[[sp]]
  cat("######", sp, " \n")
  print(tmp)
  cat(" \n\n")
  }
```
##### {-}

##### Survey size composition pearson residuals (survey 1)

```{r}
# need to separate out surveys to run this, inpN differs by survey as does n lenbins with data
# a hack
temp.surv.all <- temp.surv

temp.surv <- temp.surv %>% dplyr::filter(number ==1)

### survey, species = 1 to 11
#tiff(paste0(plotdir,"bubbleplot_survey.jpeg"), width=3000, height=2500, res=250)
ggplot(temp.surv, aes(x=year, y=lenbin, size = res_abs, color=factor(residual))) +
  geom_point(alpha=0.7) + theme(title = element_text(angle = 0, hjust = 0.5, size=15, colour="black")) +
  facet_wrap(.~species) + labs(x="year", y="length bin", title="Pearson residuals")
#dev.off()

temp.surv <- temp.surv.all
```
##### Survey size composition pearson residuals (survey 2)

```{r}
# need to separate out surveys to run this, inpN differs by survey as does n lenbins with data
# a hack
temp.surv.all <- temp.surv

temp.surv <- temp.surv %>% dplyr::filter(number ==2)

### survey, species = 1 to 11
#tiff(paste0(plotdir,"bubbleplot_survey.jpeg"), width=3000, height=2500, res=250)
ggplot(temp.surv, aes(x=year, y=lenbin, size = res_abs, color=factor(residual))) +
  geom_point(alpha=0.7) + theme(title = element_text(angle = 0, hjust = 0.5, size=15, colour="black")) +
  facet_wrap(.~species) + labs(x="year", y="length bin", title="Pearson residuals")
#dev.off()

temp.surv <- temp.surv.all
```
##### Survey OSA residuals

```{r}
 #sp <- "Atlantic_cod"
plot_osa <- list()
especies<-unique(temp.surv$species)
  for (sp in especies) {

survdat <- temp.surv %>%
  filter(number == 1, species == sp)
obs <- survdat %>% 
  select(year, lenbin, obs_value) %>% 
  mutate(obs_value = obs_value + 0.00001) %>% 
  pivot_wider(names_from = "year",
              values_from = "obs_value") %>% 
  column_to_rownames(var = "lenbin") %>% 
  as.matrix()
pred <- survdat %>% 
  select(year, lenbin, pred_value) %>% 
  pivot_wider(names_from = "year",
              values_from = "pred_value") %>% 
  column_to_rownames(var = "lenbin") %>% 
  as.matrix()
res<-resMulti(obs,pred)
plot_osa[[sp]] <- res
}
```

##### Survey 1 OSA residuals {.tabset}
```{r, results = 'asis'}
  for (sp in especies) {
  tmp <- plot_osa[[sp]]
  cat("######", sp, " \n")
  plot(tmp)
  cat(" \n\n")
  }

```
##### {-}

```{r}
#sp <- "Atlantic_cod"
plot_osa <- list()
especies<-unique(long_surv$species)
  for (sp in especies) {

survdat <- temp.surv %>%
  filter(number == 2, species == sp)
obs <- survdat %>% 
  select(year, lenbin, obs_value) %>% 
  mutate(obs_value = obs_value + 0.00001) %>% 
  pivot_wider(names_from = "year",
              values_from = "obs_value") %>% 
  column_to_rownames(var = "lenbin") %>% 
  as.matrix()
pred <- survdat %>% 
  select(year, lenbin, pred_value) %>% 
  pivot_wider(names_from = "year",
              values_from = "pred_value") %>% 
  column_to_rownames(var = "lenbin") %>% 
  as.matrix()
if(dim(pred)[1] == 0){plot_osa[[sp]] <- NULL}
else{
res<-resMulti(obs,pred)
plot_osa[[sp]] <- res
}


}
```

##### Survey 2 OSA residuals {.tabset}

```{r, results = 'asis'}
  for (sp in especies) {
  tmp <- plot_osa[[sp]]
  cat("######", sp, " \n")
  ifelse(!is.null(tmp), plot(tmp), print(tmp))
  cat(" \n\n")
  }
```
